 \documentclass[12pt]{letter}
 
\usepackage{fancyhdr}
\usepackage{lastpage}

%\usepackage{geometry}
\usepackage[paper=a4paper,textwidth=140mm,left=2.5cm,right=2.5cm,top=1.4cm,bottom=0.8cm]{geometry}

%\pagestyle{fancy}
%\lhead{\footnotesize \parbox{11cm}{Draft 1} }
%\lfoot{\footnotesize \parbox{11cm}{\textit{2}}}\cfoot{}
%\rhead{\footnotesize 3}
%\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}}
%\renewcommand{\headheight}{24pt}
%\renewcommand{\footrulewidth}{0.4pt}


\address{\phantom{a}}
       

      
\begin{document}
 
%\signature{Ileyk E\sc{l mellah}} 
 
\begin{letter} {}% Madame la Proviseure \\ Lyc\'ee  C{\sc harlemagne} \\ 14 rue Charlemagne
     % \\ 75004 {\sc paris}}
      
%\newgeometry{left=2cm,right=2cm,top=2.5cm,bottom=2.5cm}


\date{28\textsuperscript{th} March, 2016}

\opening{Dear Professors,}
 
\thispagestyle{empty}
 
%\thispagestyle{headings}
%\markright{John Smith\hfill On page styles\hfill}
 
%\thispagestyle{fancy}

\hspace*{0.5cm} I am applying to the postdoctoral position in Theory and Computation of Stellar Accretion and Winds in the Astrophysics group of the University of Exeter. I am currently a third year graduate student at the AstroParticule \& Cosmology laboratory in the University of Paris 7 Diderot. Since fall 2013, I have been conducting my PhD research in Computational Astrophysics under the supervision of Fabien Casse and Andrea Goldwurm and I am expected to defend in September 2016.\\
\hspace*{0.5cm} After my studies at the \textsc{ens}, I volunteered to join Saul Rappaport at \textsc{mit} in 2011-12. There, I contributed to his efforts to make the most of the Kepler satellite data. I used Doppler boosting to develop a photometric method to measure radial velocities in high mass ratio eclipsing stellar binaries and deduce the mass of the faint M-star. I also took part in the discovery and characterization of short-period transiting exoplanets. This inspiring insight into binary systems drove me into the study of one of their turbulent twilight, the X-ray binaries. Fabien Casse convinced me of the relevance of numerical simulations to complement the analytical skills I had acquired during the previous years. Indeed, the diversity of behaviours of those systems suggests an unavoidable need to pay attention to non-linear evolutions whose full analytical derivation remains beyond our current scope. If I have mostly got interested in winds of Supergiant stars in X-ray binairies for the last two years, I am now willing to adapt the parametrization I have set up to other systems such as young low-mass stars. The interplay between the stellar wind, rotation and magnetic activity along with the interaction with the surrounding disc offers a fruitful multi-Physics playground I am eager to explore. Thanks to new hardware technologies and optimized algorithms available, high performance \textsc{mhd} simulations can be game changers, provided we first rely on semi-analytical models to know where to look and what to look for.\\
\hspace*{0.5cm} I am applying to this postdoctoral position in Exeter for I believe my experience and commitment to computational and theoretical Astrophysics make me well qualified to meet the needs of the Theoretical Astrophysics group. I want to capitalize on the numerical expertise I have acquired and develop new numerical setups able to tackle multi-scale and multi-Physics problems. I think I can fit in an enthusiastic and stimulating environment such as the Theoretical Astrophysics Group.\\
\hspace*{0.5cm} Finally, I have passed the French Agr\'egation in Physics where I ranked second and was granted teaching responsibilities at the Paris 7 Diderot University for the last three years. I also actively took part in the organization of the \textit{Rencontre des Jeunes Physiciens} (Meeting of the Young Physicists) and in the promotion of Physics in festivals. I do intend to pursue my outreach and organizing activities and would gladly teach and monitor junior fellows.\\
\hspace*{0.5cm} I look forward to hearing from you.\\
 
Sincerely,
 
\closing{Ileyk E{\sc l Mellah}} 


  \end{letter}
  
  
 \end{document}